Constraining the Interiors of Asteroids Through Close Encounters

### 2022-04-19 paper/main.tex
Updated density adjoints

The log likelihood resulting from this uncertainty model is (excluding additive constants)
\begin{equation}
  \begin{split}
  \ln \mathcal{L} = -\frac{1}{2}\sum_{i = 0}\Bigg[&\frac{\cos^{-1} (\bm \Omega_i^* \cdot \bm \Omega_i/(\Omega_i^* \Omega_i))^2}{\sigma_\theta^2}\\
  +&\frac{\ln \parens{\Omega_i /\Omega_i^*}^2}{\sigma_\rho^2} + 2\ln\frac{\Omega_i}{\Omega_i^*}\Bigg].
  \end{split}
  \label{eqn:log-likelihood}
\end{equation}
where $\Omega_i$ is the $i$th spin vector in the data set.

@@

To propagate uncertainties from $\bm K$ to $\bm \Theta$ and therefore $\rho(\bm r)$, we need the covariance matrix $\Sigma_K$ for $\bm K$. First, assume that the hybrid frame is offset from the body-fixed frame by some small angle $\Delta \gamma$, which results from uncertainty in $\gamma_0$. Then
\begin{equation}
  K_{\ell m}^\mathrm{hybrid} = e^{-im\Delta \gamma}K_{\ell m}^\mathrm{body-fixed}.
  \label{eqn:body-fixed-to-hybrid}
\end{equation}
by equation \ref{eqn:ylm-rotation}. Since $K_{\ell m}^\mathrm{body-fixed}$ was obtained by an MCMC fit, a large set of PPD-distributed samples is available for $\Delta \gamma$ and $K_{\ell m}^\mathrm{body-fixed}$, and the covariance matrix $\Sigma_K$ can be computed statistically in the hybrid frame by applying equation \ref{eqn:body-fixed-to-hybrid} to the samples. Then, propagation of uncertainty guarantees that the covariance matrix of $\bm \Theta$ is $\Sigma_\Theta = A \Sigma_K A^T$ and the density distribution and uncertainty on density distribution are equal to
\begin{equation}
  \rho(\bm r) = \bm B(\bm r)^T A\bm K \qquad \sigma^2_\rho(\bm r) = \bm B(\bm r)^T A \Sigma_K A^T \bm B(\bm r).
  \label{eqn:unc-rho}
\end{equation}

@@

\begin{figure*}
  \centering
  \includegraphics[width=0.7\textwidth]{figs/example-residuals.pdf}
  \caption{Data, best fit results, and residuals for a fit to synthetic data simulated for an asymmetric reference asteroid. The standard deviation of the data is plotted as an uncertainty band in the residuals plot, but these do not capture covariance between data points.}
  \label{fig:example-residuals}
\end{figure*}

@@

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{figs/example-corner.pdf}
  \caption{A corner plot for the fit results ten parameters to synthetic data simulated for an asymmetric reference asteroid. Marginal posterior PDFs for each parameter are shown (histograms) along with two-dimensional PDFs (contours). Individual points in the contours are samples from the MCMC fit, and the contours enclose 1, 2, and 3$\sigma$ of them. True values (which are zero here due to the parameters being displayed relative to the true values) are also shown (blue lines).}
  \label{fig:example-corner}
\end{figure*}

@@

In this section, we discuss the physical properties of the asteroid encounter system (section \ref{sec:spin-evolution}), as well as the sensitivity of the density moment uncertainty to encounter properties (section \ref{sec:fit-uncertainty}), together with an analysis of advantages and disadvantages Jupiter encounters over Earth (section \ref{sec:jupiter-earth})Finally, we discuss the kinds of distributions produced by the density distribution extraction methods (section \ref{sec:density-compare}).

@@

\begin{figure*}
  \centering
  \includegraphics[width=0.7\linewidth]{figs/nominal-data-sym.pdf}
  \includegraphics[width=0.7\linewidth]{figs/nominal-data-asym.pdf}
  \caption{Angular velocity data simulated for the symmetric (above) and asymmetric (below) reference asteroids. The true angular velocity evolution is shown as a black line. Also plotted is the deviation of the data for PPD-distributed perturbations to the asteroid shape (bands). Bands contain 68.3\%, 95.5\%, and 99.7\% of the 1000 simulations run.}
  \label{fig:example-data}
\end{figure*}

@@

\afterpage{
  \begin{landscape}
    \begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figs/scan-all-1.pdf}
    \caption{1 and 2$\sigma$ confidence intervals for the first-order parameter PPDs (\textit{top}) and second-order parameters (\textit{bottom}) as a function of various encounter parameters (\textit{columns}). The vertical dashed line indicates the reference asteroid values of the parameters. The red vertical lines indicate when $\sigma =0.01$.}
    \label{fig:scan-perigee}
    \label{fig:scan-vex}
    \label{fig:scan-theta}
    \label{fig:scan-rho}
    \end{figure}
  \end{landscape}
}

\afterpage{
  \begin{landscape}
    \begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figs/scan-all-2.pdf}
    \caption{1 and 2$\sigma$ confidence intervals for the first-order parameter PPDs (\textit{top}) and second-order parameters (\textit{bottom}) as a function of various encounter parameters (\textit{columns}). The vertical dashed line indicates the reference asteroid values of the parameters. The red vertical lines indicate when $\sigma =0.01$.}
    \label{fig:scan-am}
    \label{fig:scan-cadence}
    \label{fig:scan-period}
    \label{fig:scan-gap}
    \end{figure}
  \end{landscape}
}

@@

In figure \ref{fig:scan-spin}, we display 1$\sigma$ uncertainties for all parameters as a function of the direction of $\bm \Omega_0$, mapped onto the unit sphere in the inertial frame. Our samples for $\bm \Omega_0$ were laid out on a Fibonacci sphere to ensure they were roughly evenly spaced (marked in figure \ref{fig:scan-spin-avg}). To highlight common features across the parameters, we also display the average 1$\sigma$ sensitivity in figure \ref{fig:scan-spin-avg}. The average is weighted such that the uncertainty map for each parameter contributes an equal amount
(the weight of each map is set to one-tenth of the map's mean). This average map is presented in two different projections to allow data at $\unit Z$ to be read.

\begin{figure*}
  \centering
  \includegraphics[width=0.8\textwidth]{figs/spin-pole.pdf}
  \caption{$1\sigma$ uncertainties for the first-order parameters (\textit{top}) and second-order (\textit{bottom}) as a function of the initial direction of spin in the inertial frame. All maps are made in the Mollweide projection. The orange star indicates the reference spin pole. The red contours enclose regions where $\sigma \geq 0.01$.}
  \label{fig:scan-spin}
\end{figure*}

\begin{figure*}
  \centering
  \includegraphics[width=0.3\textwidth]{figs/spin-pole-avg-mollweide.pdf}
  \includegraphics[width=0.3\textwidth]{figs/spin-pole-avg-polar.pdf}
  \caption{The weighted average of the uncertainties shown in figure \ref{fig:scan-spin}, in Mollweide (\textit{left}) and polar (\textit{right}) projections in the inertial frame. See text for a description of how the average was computed. Black dots indicate the Fibonacci-sphere-distributed locations of sample spin poles, and the orange star indicates the reference spin pole. The polar projection is centred at the south pole, or $-\unit Z$.}
  \label{fig:scan-spin-avg}
\end{figure*}

>>>>

The log likelihood resulting from this uncertainty model is (excluding additive constants)
\begin{equation}
  \begin{split}
  \ln \mathcal{L} = -\frac{1}{2}\sum_{i = 0}\Bigg[&\frac{\cos^{-1} (\bm \Omega_i^* \cdot \bm \Omega_i/(\Omega_i^* \Omega_i))^2}{\sigma_\theta^2}\\
  &+\frac{\ln \parens{\Omega_i /\Omega_i^*}^2}{\sigma_\rho^2} + 2\ln\frac{\Omega_i}{\Omega_i^*}\Bigg].
  \end{split}
  \label{eqn:log-likelihood}
\end{equation}
where $\Omega_i$ is the $i$th spin vector in the data set.

@@

To propagate uncertainties from $\bm K$ to $\bm \Theta$ and therefore $\rho(\bm r)$, we need the covariance matrix $\Sigma_K$ for $\bm K$. First, assume that the hybrid frame is offset from the body-fixed frame by some small angle $\Delta \gamma$, which results from uncertainty in $\gamma_0$. Then
\begin{equation}
  K_{\ell m}^\mathrm{hybrid} = e^{-im\Delta \gamma}K_{\ell m}^\mathrm{body-fixed}.
  \label{eqn:body-fixed-to-hybrid}
\end{equation}
by equation \ref{eqn:ylm-rotation}. Since $K_{\ell m}^\mathrm{body-fixed}$ was obtained by an MCMC fit, a large set of PPD-distributed samples is available for $\Delta \gamma$ and $K_{\ell m}^\mathrm{body-fixed}$, and the covariance matrix $\Sigma_K$ can be computed statistically in the hybrid frame by applying equation \ref{eqn:body-fixed-to-hybrid} to the samples. Then, propagation of uncertainty guarantees that the covariance matrix of $\bm \Theta$ is $\Sigma_\Theta = A \Sigma_K A^*$ and the density distribution and uncertainty on density distribution are equal to
\begin{equation}
  \rho(\bm r) = \bm B(\bm r)^* A\bm K \qquad \sigma^2_\rho(\bm r) = \bm B(\bm r)^* A \Sigma_K A^* \bm B(\bm r).
  \label{eqn:unc-rho}
\end{equation}
where $A^*$ is the adjoint of $A$.

@@

\begin{figure*}
  \centering
  \includegraphics[width=0.6\textwidth]{figs/example-residuals.pdf}
  \caption{Data, best fit results, and residuals for a fit to synthetic data simulated for an asymmetric reference asteroid. Uncertainty bands are also shown.}
  \label{fig:example-residuals}
\end{figure*}

\begin{figure*}
  \centering
  \includegraphics[width=0.85\textwidth]{figs/example-corner.png}
  \caption{PPDs extracted from synthetic encounter data for the asymmetric reference asteroid. Samples from the MCMC fit are shown as individual points, and the contours enclose 1, 2, and 3$\sigma$ confidence regions. True values are shown as blue lines.}
  \label{fig:example-corner}
\end{figure*}

@@

@@

In this section, we discuss the physical properties of the asteroid encounter system (section \ref{sec:spin-evolution}), as well as the sensitivity of the density moment uncertainty to encounter properties (section \ref{sec:fit-uncertainty}), together with an analysis of advantages and disadvantages Jupiter encounters over Earth (section \ref{sec:jupiter-earth}). Finally, we discuss the kinds of distributions produced by the density distribution extraction methods (section \ref{sec:density-compare}).

@@

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figs/nominal-data-sym.pdf}
  \includegraphics[width=\linewidth]{figs/nominal-data-asym.pdf}
  \caption{Angular velocity data simulated for the symmetric (above) and asymmetric (below) reference asteroids. The true angular velocity evolution is shown as a black line. Also plotted is the deviation of the data for PPD-distributed perturbations to the asteroid shape (bands). Bands contain 68.3\%, 95.5\%, and 99.7\% of the 1000 simulations run.}
  \label{fig:example-data}
\end{figure}

@@

\afterpage{
  \begin{landscape}
    \begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figs/scan-all.pdf}
    \caption{1 and 2$\sigma$ confidence intervals for the first-order parameter PPDs (\textit{top}) and second-order parameters (\textit{bottom}) as a function of various encounter parameters (\textit{columns}). The vertical dashed line indicates the reference asteroid values of the parameters. The red vertical lines indicate when $\sigma =0.01$.}
    \label{fig:scan-perigee}
    \label{fig:scan-vex}
    \label{fig:scan-theta}
    \label{fig:scan-rho}
    \end{figure}
  \end{landscape}
}

\afterpage{
  \begin{landscape}
    \begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figs/scan-all.pdf}
    \caption{1 and 2$\sigma$ confidence intervals for the first-order parameter PPDs (\textit{top}) and second-order parameters (\textit{bottom}) as a function of various encounter parameters (\textit{columns}). The vertical dashed line indicates the reference asteroid values of the parameters. The red vertical lines indicate when $\sigma =0.01$.}
    \label{fig:scan-am}
    \label{fig:scan-cadence}
    \label{fig:scan-period}
    \label{fig:scan-gap}
    \end{figure}
  \end{landscape}
}

@@

In figure \ref{fig:scan-spin}, we display 1$\sigma$ uncertainties for all parameters as a function of the direction of $\bm \Omega_0$, mapped onto the unit sphere in the inertial frame. Our samples for $\bm \Omega_0$ were laid out on a Fibonacci sphere to ensure they were roughly evenly spaced (marked in figure \ref{fig:scan-spin-avg}). To highlight common features across the parameters, we also display the average 1$\sigma$ sensitivity in figure \ref{fig:scan-spin-avg}. The average is weighted such that the uncertainty map for each parameter contributes an equal amount
(the weight of each map is set to one-tenth of the map's mean). This average map is presented in two different projections to allow data at $\unit Z$ to be read.
`'
\begin{figure*}
  \centering
  \includegraphics[width=0.8\textwidth]{figs/spin-pole.png}
  \caption{$1\sigma$ uncertainties for the first-order parameters (\textit{top}) and second-order (\textit{bottom}) as a function of the initial direction of spin in the inertial frame. All maps are made in the Mollweide projection. The orange star indicates the reference spin pole. The red contours enclose regions where $\sigma \geq 0.01$.}
  \label{fig:scan-spin}
\end{figure*}

\begin{figure*}
  \centering
  \includegraphics[width=0.3\textwidth]{figs/spin-pole-avg-mollweide.png}
  \includegraphics[width=0.3\textwidth]{figs/spin-pole-avg-polar.png}
  \caption{The weighted average of the uncertainties shown in figure \ref{fig:scan-spin}, in Mollweide (\textit{left}) and polar (\textit{right}) projections in the inertial frame. See text for a description of how the average was computed. Black dots indicate the Fibonacci-sphere-distributed locations of sample spin poles, and the orange star indicates the reference spin pole. The polar projection is centred at the south pole, or $-\unit Z$.}
  \label{fig:scan-spin-avg}
\end{figure*}
